### 2022-11-30 paper_OJA.tex
reduce Appendix

\PassOptionsToPackage{usenames,dvipsnames}{xcolour}
\documentclass[twocolumn,twocolappendix,nofootinbib,iop]{openjournal}

% Add your own macros here:
\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amssymb,amstext}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage{ae,aecompl}
\usepackage[utf8]{inputenc}
%\usepackage{hyperref}
\usepackage[colorlinks,allcolors=blue]{hyperref}
\usepackage[figure,figure*]{hypcap}
\usepackage{natbib}
\usepackage{url}
\usepackage{mdwlist}
\usepackage{multirow}
\urlstyle{same}
\usepackage{lineno}
\usepackage{fontawesome}

% jZ had to remove this to avoid clash
% \usepackage[
% singlelinecheck=false % <-- important
% ]{caption}

%to cross out a sentence: \st{Hello World}
\usepackage{soul}

%%%% usepackage{minted}
\usepackage{xcolor}
\definecolor{maroon}{cmyk}{0, 0.87, 0.68, 0.32}
\definecolor{halfgray}{gray}{0.55}
\definecolor{ipython_frame}{RGB}{207, 207, 207}
\definecolor{ipython_bg}{RGB}{247, 247, 247}
\definecolor{ipython_red}{RGB}{186, 33, 33}
\definecolor{ipython_green}{RGB}{0, 128, 0}
\definecolor{ipython_cyan}{RGB}{64, 128, 128}
\definecolor{ipython_purple}{RGB}{170, 34, 255}

\usepackage{listings}

\lstset{
    breaklines=true,
    %
    extendedchars=true,
    literate=
    {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
    {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
    {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
    {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
    {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
    {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
    {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
    {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
    {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
    {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
    {€}{{\EUR}}1 {£}{{\pounds}}1
}

%%
%% Python definition (c) 1998 Michael Weber
%% Additional definitions (2013) Alexis Dimitriadis
%% modified by me (should not have empty lines)
%%
\lstdefinelanguage{iPython}{
    morekeywords={access,and,break,class,continue,def,del,elif,else,except,exec,finally,for,from,global,if,import,in,is,lambda,not,or,pass,print,raise,return,try,while},%
    %
    % Built-ins
    morekeywords=[2]{abs,all,any,basestring,bin,bool,bytearray,callable,chr,classmethod,cmp,compile,complex,delattr,dict,dir,divmod,enumerate,eval,execfile,file,filter,float,format,frozenset,getattr,globals,hasattr,hash,help,hex,id,input,int,isinstance,issubclass,iter,len,list,locals,long,map,max,memoryview,min,next,object,oct,open,ord,pow,property,range,raw_input,reduce,reload,repr,reversed,round,set,setattr,slice,sorted,staticmethod,str,sum,super,tuple,type,unichr,unicode,vars,xrange,zip,apply,buffer,coerce,intern},%
    %
    sensitive=true,%
    morecomment=[l]\#,%
    morestring=[b]',%
    morestring=[b]",%
    %
    morestring=[s]{'''}{'''},% used for documentation text (mulitiline strings)
    morestring=[s]{"""}{"""},% added by Philipp Matthias Hahn
    %
    morestring=[s]{r'}{'},% `raw' strings
    morestring=[s]{r"}{"},%
    morestring=[s]{r'''}{'''},%
    morestring=[s]{r"""}{"""},%
    morestring=[s]{u'}{'},% unicode strings
    morestring=[s]{u"}{"},%
    morestring=[s]{u'''}{'''},%
    morestring=[s]{u"""}{"""},%
    %
    % {replace}{replacement}{lenght of replace}
    % *{-}{-}{1} will not replace in comments and so on
    literate=
    {á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'i}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
    {Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
    {à}{{\`a}}1 {è}{{\`e}}1 {ì}{{\`i}}1 {ò}{{\`o}}1 {ù}{{\`u}}1
    {À}{{\`A}}1 {È}{{\'E}}1 {Ì}{{\`I}}1 {Ò}{{\`O}}1 {Ù}{{\`U}}1
    {ä}{{\"a}}1 {ë}{{\"e}}1 {ï}{{\"i}}1 {ö}{{\"o}}1 {ü}{{\"u}}1
    {Ä}{{\"A}}1 {Ë}{{\"E}}1 {Ï}{{\"I}}1 {Ö}{{\"O}}1 {Ü}{{\"U}}1
    {â}{{\^a}}1 {ê}{{\^e}}1 {î}{{\^i}}1 {ô}{{\^o}}1 {û}{{\^u}}1
    {Â}{{\^A}}1 {Ê}{{\^E}}1 {Î}{{\^I}}1 {Ô}{{\^O}}1 {Û}{{\^U}}1
    {œ}{{\oe}}1 {Œ}{{\OE}}1 {æ}{{\ae}}1 {Æ}{{\AE}}1 {ß}{{\ss}}1
    {ç}{{\c c}}1 {Ç}{{\c C}}1 {ø}{{\o}}1 {å}{{\r a}}1 {Å}{{\r A}}1
    {€}{{\EUR}}1 {£}{{\pounds}}1
    %
    {^}{{{\color{ipython_purple}\^{}}}}1
    {=}{{{\color{ipython_purple}=}}}1
    %
    {+}{{{\color{ipython_purple}+}}}1
    {-}{{{\color{ipython_purple}-}}}1
    {*}{{{\color{ipython_purple}$^\ast$}}}1
    {/}{{{\color{ipython_purple}/}}}1
    %
    {+=}{{{+=}}}1
    {-=}{{{-=}}}1
    {*=}{{{$^\ast$=}}}1
    {/=}{{{/=}}}1,
    literate=
    *{-}{{{\color{ipython_purple}-}}}1
     {?}{{{\color{ipython_purple}?}}}1,
    %
    identifierstyle=\color{black}\ttfamily,
    commentstyle=\color{ipython_cyan}\ttfamily,
    stringstyle=\color{ipython_red}\ttfamily,
    keepspaces=true,
    showspaces=false,
    showstringspaces=false,
    %
    rulecolor=\color{ipython_frame},
    frameround={t}{t}{t}{t},
    numbers=none,
    numberstyle=\tiny\color{halfgray},
    %
    %
    backgroundcolor=\color{ipython_bg},
    %   extendedchars=true,
    %basicstyle=\scriptsize,
    basicstyle=\ttfamily\footnotesize,
    columns=fullflexible,
    keywordstyle=\color{ipython_green}\ttfamily,
}

\newcommand{\nblink}[1]{\href{https://github.com/DifferentiableUniverseInitiative/jax-cosmo-paper/blob/master/notebooks/#1.ipynb}{\faFileCodeO}}
\newcommand{\github}{\href{https://github.com/DifferentiableUniverseInitiative/jax\_cosmo}{\faGithub}}

%%% color Comments 
%% JEC : \FL produces a clash we another definition (but coannot figure out which one)  
\newcommand{\FrL}[1]{{\color{cyan}FL: #1}}
\newcommand{\JZ}[1]{{\color{purple}JZ: #1}}
\newcommand{\JEC}[1]{{\color{magenta}JEC: #1}}

%%%% new command (JEC)
\newcommand{\nn}{\nonumber}
\newcommand{\numpyro}{\texttt{NumPyro}}
\newcommand{\bydef}{:=}
\newcommand{\jaxcosmo}{\texttt{jax-cosmo}}
\newcommand{\autodiff}{\texttt{autodiff}}

\begin{document}
\journalinfo{The Open Journal of Astrophysics}
%\submitted{submitted xxx, revised yyy, accepted zzz}

\title{\jaxcosmo: An End-to-End Differentiable and GPU Accelerated Cosmology Library}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{
J.~E. Campagne$^{1,\ast}$, 
F. Lanusse$^2$, 
J. Zuntz$^3$,\\
\jaxcosmo\ contributors}
\thanks{$^\ast$jean-eric.campagne@ijclab.in2p3.fr}

\affiliation{$^1$Université Paris-Saclay, CNRS/IN2P3, IJCLab, 91405 Orsay, France}

%\date{\today}

\begin{abstract}
The recent JAX library has created a new ecosystem from which probabilistic programming software can take enormous benefit: batch acceleration, just-in-time compilation, and automatic optimization of code for different hardware modalities (CPU, GPU, TPU) can provide huge speed-ups to a wide range of different problems. In particular, JAX exposes an \textit{automatic differentiation} mechanism, which can take a python function that uses JAX and return its (vector) derivative.  This gives us the opportunity to apply a range of powerful but otherwise unfeasible algorithms used in Bayesian inference, such as Hamiltonian Monte Carlo (HMC) and Variational Inference.

To take advantage of these possibilities within cosmology we have developed the \jaxcosmo\ library, which implements background evolution, linear and non linear power spectra (using \texttt{halofit} or the Eisenstein and Hu transfer function), as well as angular power spectra ($C_\ell$) with the Limber approximation for galaxy and weak lensing probes. We discuss algorithms made possible by this library, and present comparisons with the Core Cosmology Library as a benchmark, and run a series of tests using the Dark Energy Survey Year 1 3x2pt analysis with the \numpyro\ library to demonstrate practical usage.

We show that clear improvements are possible using HMC compared to Metropolis-Hasting, and that the Normalizing Flows using the Neural Transport is a promising methodology.
%{\color{red}
%;in the early phases of Bayesian model development JAZ not sure what this means so I left it out}

\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

%% \linenumbers

%% main text

\section{Introduction}
Bayesian inference has been widely used in cosmology in the form of Monte Carlo Markov Chains (MCMC) since the work of \citep{2001ApJ...563L..95K,2003MNRAS.341.1084R} and has been the keystone for past and current analysis thanks partly to packages as  \texttt{CosmoMC} \citep{2002PhRvD..66j3511L}, \texttt{CosmoSIS} \citep{2015A&C....12...45Z}, \texttt{MontePython} \citep{2019PDU....24..260B} as well as   \texttt{Cobaya} \citep{2019ascl.soft10019T,2021JCAP...05..057T}; see, for instance, the list of citations to these popular packages for an idea of the wide usage in the community.

% \footnote{\url{https://cosmologist.info/cosmomc/readme.html}}
% \footnote{\url{http://bitbucket.org/joezuntz/cosmosis}}
% \footnote{\url{https://cobaya.readthedocs.io}}
% \footnote{\url{https://github.com/brinckmann/montepython_public}}

% Note that some direct optimization of likelihood function has also been successfully undertaken for instance for some Planck analysis \citep{2014A&A...566A..54P}, 
Since the development of these MCMC packages, major advances have been made in automatic differentiation (\textit{autodiff}) \citep{10.5555/3122009.3242010, Margossian2019}, a set of technologies for transforming pieces of code into their derivatives.

While these tools have especially been applied to neural network optimization and machine learning in general, they can also enable classical statistical methods that require the derivatives of (e.g. likelihood) functions to operate: we consider such methods in this paper. \textit{Autodiff} has been implemented in widely used libraries like \texttt{Stan} \citep{JSSv076i01}, \texttt{TensorFlow}  \citep{tensorflow2015-whitepaper}, and  \texttt{PyTorch} \citep{NEURIPS2019_9015}.

A recent entrant to this field is the \texttt{JAX} library\footnote{\url{https://jax.readthedocs.io}} \citep{jax2018github} which has undergone rapid development and can automatically differentiate native \texttt{Python} and \texttt{NumPy} functions, offering a speed up to the development process and indeed code runtimes. \texttt{JAX} offers an easy parallelization mechanism (\texttt{vmap}), and just-in-time compilation (\texttt{jit}) and optimization targetting CPU, GPU and TPU hardware thanks to the \texttt{XLA} library that powers TensorFlow. These attractive features have driven wide adoption of \texttt{JAX} in computational research, and motivate us to consider its usage in cosmology.

\texttt{JAX} contains bespoke reimplementation of packages such as \texttt{jax.numpy} and \texttt{jax.scipy}, as well as example libraries as \texttt{Stax} for simple but flexible neural network development. It has formed the seed for a wider ecosystem of packages, including, for example: 
\texttt{Flax} \citep{flax2020github} a high-performance neural network library, \texttt{JAXopt} \citep{jaxopt_implicit_diff} an hardware accelerated, batchable and differentiable collection of optimizers, \texttt{Optax} \citep{optax2020github} a gradient processing and optimization library, and \numpyro\ \citep{phan2019composable,bingham2019pyro}, a probabilistic programming language (PPL) that is used in this paper. Other PPL packages such as \texttt{PyMC} \citep{Salvatier2016} have switched to JAX backend in recent versions\footnote{A more exhaustive list and rapidly growing list of packages can be found at \url{https://project-awesome.org/n2cholas/awesome-jax}}.

% I don't think we need all these URLs, it's getting excessive, especially when they have papers:
%\footnote{\url{https://flax.readthedocs.io/}}
% \footnote{\url{https://jaxopt.github.io}}
% \footnote{\url{https://optax.readthedocs.io}}
% \footnote{\url{https://num.pyro.ai}}
% \footnote{\url{https://www.pymc.io/projects/docs/en/stable/installation.html}}
% \footnote{\url{https://www.tensorflow.org/xla}}
% \footnote{\url{https://www.tensorflow.org/}}
% \footnote{\url{https://pytorch.org/docs/}}
% \footnote{\url{https://mc-stan.org/users/documentation}}

In this context we have developed the open source \jaxcosmo\ library\footnote{\url{https://github.com/DifferentiableUniverseInitiative/jax_cosmo}}, which we present in this paper. The package represents a first step in making the powerful features described above useful for cosmology; it implements a selection of theory predictions for key cosmology observables as differentiable JAX functions.

We give an overview of the code's design and contents in Section~\ref{sec-jaxcosmo-design}. We show how to use it for rapid and numerically stable Fisher forecasts in Section~\ref{sec-fisher-forecast}, in more complete gradient-based cosmological inference with variants of Hamiltonian Monte Carlo and its variant No-U-Turn, and ML-accelerated Stochastic Variational Inference in Section~\ref{sec:chmc}. We discuss and compare these methods in Section~\ref{sec-results} and conclude in Section~\ref{sec-conclusion}.

\section{JAX: Autograd meets XLA}
\label{sec-primer}

The aim of this session is to provide a brief technical primer on JAX, necessary to fully grasp the potential of a cosmology library implemented in this framework.

% Maybe a short history explaining what autograd and XLA are.
\FrL{Add a short history of XLA and Autograd.}

\FrL{Is this too much information below? On the one hand I don't want to go too technical, on the other hand, we talk about stuff like jax.grad/hessian, JIT, and vmap in the rest of the paper, it feels like these concepts should be introduced somewhere first.}

\paragraph{Automatic Differentiation} Traditionally, two different approaches have been used in cosmology to obtain derivatives of given numerical expressions.

The first approach is to derive analytically the formula for the derivatives of interest \citep{2019A&A...625A..64J}%JEC: This reference should be validated by F.L
, with or without the help of tools such as \texttt{Mathematica}\footnote{\url{https://www.wolfram.com/mathematica}.}. This is however only practical typically for simple analytical models.

The second approach is to compute numerical derivatives by finite differences. This approach can be applied on black-box models of abitrary complexity (from Boltzmann codes \FrL{some citations} to cosmological simulations \FrL{some citations}). However it is notoriously difficult to obtain stable derivatives by finite differences \FrL{cite papers trying to do fisher}. In addition, their computational cost does not scale well with the number of parameters (a minimum of $(2N+1)$ model evaluations is typically required for $N$ parameters), making them completely impractical whenever derivatives are needed on the fly.\FrL{this last sentence is not great.}

Automatic differentiation frameworks like JAX take a different approach. They trace the execution of a given model and decompose this trace into fundamental operations with known derivatives (e.g. the derivative of a multiplication operation is known). Then, by applying the chain rule formula, the computational graph for the derivatives (of any order) of the model can be built from the known derivatives of every elementary operations. A new function corresponding to the derivative of the original function is therefore built automatically for the user. We direct the interested reader to \FrL{a good reference} for an in-depth introduction to automatic differentiation.

JAX provides in particular a number of operators (\texttt{jax.grad}, \texttt{jax.jacobian}, \texttt{jax.hessian}) which can compute derivatives of any function written in JAX.
\begin{lstlisting}[language=iPython]
# Let us define a simple function 
def f(x):
	return y = 5 * x + 2
# Take the derivative
df_dx = jax.grad(f)
# df_dx is a new function that always return 5 
\end{lstlisting}

\textbf{Why is that interesting?} \autodiff\ makes it possible to obtain \textit{exact gradients of cosmological likelihoods} with respect to all input parameters at the cost of only 2 likelihood evaluations.

\paragraph{Just In Time Compilation (JIT)} Despite its convenience and wide adoption in astrophysics, Python still suffers from slow execution times compared to fully compiled languages such as C/C++. One approach to mitigate these issues and make Python code fast is Just In Time compilation, which traces the execution of a given Python function the first time it is called, and compiles it into a fast executable (by-passing the Python interpreter), which can be transparently used in subsequent calls to this function. Compared to other strategies for speeding up Python code \FrL{cpython and such}, JIT allows the user to simply write plain Python code, and reap \FrL{spell check} the benefits of compiled code.

A number of libraries allowing for JIT have already been used in astrophysics, in particular Numba (ngmix), and \FrL{The JIT code developped by our friends in Zurich}. JAX stands out compared to these other frameworks in that it relies on the highly optimized XLA library \FrL{citation} for executing the compiled expressions. XLA is continuously developed by Google as part of the TensorFlow project, for efficient training and inference of large scale deep learning applications, and as such supports computational backends such as GPU an TPU clusters.
\FrL{Maybe we want to say why GPU is such a good idea for our use cases.}

In JAX, jitting is achieving simply by transforming a function with \texttt{jax.jit} operation:
\begin{lstlisting}[language=iPython]
# Let us redefine our function
def f(x):
	return y = 5 * x + 2
# And JIT it
jitted_f = jax.jit(f)
# The first call to jitted_f will be relatively slow
# subsequent calls will be extremely fast
# and run as a compiled code directly on GPU
\end{lstlisting}

\textbf{Why is that interesting?} JIT makes it possible to execute entire cosmological \textit{MCMC chains directly on GPUs as compiled code}, with orders of magnitude in speedup over Python code.

\paragraph{Automatic Vectorization} Another extremely powerful feature of JAX is its ability to automatically vectorize or paralellize any function. Through the same tracing mechanism used for automatic differentiation, JAX can decompose a given computation into fundamental operations and add a new \textit{batch} dimension so that the computation can be applied on a batch of inputs as opposed to individual inputs. In doing so, note that the computation will not run sequentially over all entries of the batch, but trully in parallel making full use of the intrinsic parallel architecture of modern GPUs.

In JAX automatic vectorization is achieved using the \textit{jax.vmap} operation:
\begin{lstlisting}[language=iPython]
# Our function f only applies to scalars
def f(x):
	return y = 5 * x + 2
# Applying automatic vectorization
batched_f = jax.vmap(f)
# batched_f now applies to 1D arrays
\end{lstlisting}
Again, we stress that in this example, \texttt{batched\_f} will not be implemented in terms of a for loop, but with operations over vectors. The function above is trivial, but the same approach can be used to parallelize any function, from Limber integrals, to an entire likelihood. When considering multi-devices use-case (eg. GPU or TPU), JAX provides \texttt{pmap} which compile and execute in parallel replicates of the same code on each device. Moreover, recent experimental developments deal with parallelization of function over supercomputer-scale hardware meshes. Notice that in the examples detailed in this article, we have only relied on \texttt{vmap} functionality.

\textbf{Why is that interesting?} Automatic Vectorization makes it possible to trivially parallelize cosmological likelihood evaluations, to run many parallel MCMC chains on a single GPU.

\paragraph{NumPy API compliance} Finally, the last point to note about JAX, is that it mirrors the NumPy API (with only a few exceptions). This means in practice that existing NumPy code can easily be converted to JAX. This is in contrast to other similar frameworks like TensorFlow, PyTorch, or Julia which all require the user to learn, and adapt their code to, a new API or even a new language.

\textbf{Why is that interesting?} NumPy compliance implies improved maintainability and lower barrier to entry for new contributors.

\section{Design of the \jaxcosmo\ library}
\label{sec-jaxcosmo-design}

In this section, we describe the cosmological modeling provided by \jaxcosmo, and its implementation in JAX. The general design follows that of the Core Cosmology Library (CCL) \citep{2019ApJS..242....2C}, though, in its initial release,
 \jaxcosmo\ only implements a subset of CCL features and options.

All \jaxcosmo\ data structures are organized as JAX \textit{pyTree} container objects, ensuring that two key JAX features are available to them: \textit{autodiff} and \textit{vmap}. The vmap feature enables any operation defined in JAX (including complicated composite operations) to be applied efficiently as a vector operation. The autodiff feature further makes it possible to take the derivative of any operation, automatically transforming a function that takes $n$ inputs and produces $m$ outputs into a new function that generates an $m \times n$ matrix of partial derivatives.

\jaxcosmo\ implements, for example, Runge-Kutta solvers for ODEs, and Simpson \& Romberg integration routines through this framework, so that we can automatically compute the derivative of their solutions with respect to their inputs. This includes the cosmological parameters (the standard set of $w_0 w_a CDM$ cosmological parameters is exposed, using $\sigma_8$ as an amplitude parameter), but also the other input quantities such as redshift distributions described below.

In the rest of this section we describe the cosmological calculations that are implemented using these facilities.

\subsection{Background cosmology}

The computation of the evolution of the cosmological background follows a typical implementation of a Friedmann equation (\citealp[see e.g.][]{2005A&A...443..819P}):
\begin{equation}
    E^2(a) = \frac{\mathrm{H}^2(a)}{\mathrm{H}^2_0} = \Omega_m a^{-3} + \Omega_k a^{-2} + \Omega_{de} e^{f(a)}
\end{equation}
with $a=1/(1+z)$ the scale factor related to the redshift $z$, $\mathrm{H}(a)=\dot{a}/a$ the Hubble parameter with $\mathrm{H}_0$ its present day value, $\Omega_m= \Omega_{cdm}+\Omega_b$, $\Omega_{de}=1-\Omega_k-\Omega_m$, and 
\begin{equation}
    f(a) = -3 (1 + w_0 + w_a) \ln(a) + 3 w_a (a - 1)
\end{equation}
Notably, the relativistic contributions of massless neutrinos and photon radiation, as well as the massive neutrino contribution are currently neglected. From these expressions, in \texttt{jc.background}, are computed the different cosmological distance functions, such as the radial comoving distance
\begin{equation}
     \chi(a) =  R_H \int_a^1 \frac{da^\prime}{{a^\prime}^2 E(a^\prime)}
     \label{eq:radial_comoving}
\end{equation}
with $R_H$ the Hubble radius. 
%Using the scale factor ($a$) redshift ($z$) relationship, $\chi$ can be viewed as a function of $z$. 
%
\subsection{Growth of perturbations}
%
Currently, \jaxcosmo\  implements the \citet{Eisenstein_1998} transfer function $T$ which transforms the primordial matter power spectrum to its late-time non-linear value:
\begin{equation}
    P(k, z) = P(k, z=\infty) \cdot T(k, z; \Omega_m, \Omega_b, ...),
\end{equation}
through the \textit{halofit} model by \cite{2012ApJ...761..152T} or \cite{2003MNRAS.341.1311S} without the neutrino contribution introduced by \cite{10.1111/j.1365-2966.2011.20222.x}. No Baryon feedback modeling is considered yet.

The primordial power spectrum is modelled with the standard form:
\begin{equation}
    P(k) = A k^{n_s}.
\end{equation}

The normalisation $A$ is parametrized via $\sigma_8$ at $z=0$ as 
\begin{equation}
    A = \sigma_8^2 \times \left(\frac{1}{2 \pi^2} \int_0^\infty \frac{dk}{k} k^3 P(k) W^2(kR_8) \right)^{-1}
\end{equation}
with $R_8 = 8 \mathrm{Mpc}/h$ and $W(x)$ related to the $j_1$ spherical Bessel function as
\begin{equation}
    W(x) = \frac{3j_1(x)}{x}
\end{equation}

%Future version of the library would offer the possibility to call a JAX emulator of the Cosmic Linear Anisotropy Solving System \citep{2011JCAP...07..034B}. \JZ{How close is this to being usable? If it's just an idea or an early prototype then skip this paragraph.}
%
\subsection{Angular power spectra}
%
\jaxcosmo\ is currently focused on predicting projected 2D Fourier-space 2pt galaxy lensing, clustering, and cross correlations, the $C_\ell$ angular power spectra that are a primary target of upcoming photometric surveys. The details of the implementation is in \texttt{jc.angular\_cl} which deals with the mean and Gaussian covariance matrix computations.

The angular power spectra $C_\ell^{ij}$ for the probes $(i,j)$ and for redshift bin window selections are computed in the first order Limber approximation \citep{PhysRevD.78.123506}:
\begin{align}
    C_\ell^{i,j} \approx \left(\ell+\frac{1}{2}\right)^{m_i+m_j}\int\frac{d\chi}{c^2\chi^2}K_i(\chi)K_j(\chi)\,P\left(k=\frac{\ell+1/2}{\chi},z\right),\label{eq:Cell_limber}
\end{align}
The factors $m_i$ are equal to $(0,-2)$ for the galaxy clustering and weak lensing respectively, and the functions $K(z)$ each represent a single tomographic redshift bin's number density. These tracers are implemented as two kernel functions:

\begin{description}
\item[\texttt{NumberCounts}]
    \begin{equation}
        K_i(z) = n_i(z)\ b(z)\ H(z)
    \end{equation}
where $n_i(s)$ is the redshift distribution of the sample (eg. \texttt{jc.redshift.kde\_nz} function), $b(z)$ a galaxy bias function (see  \texttt{jc.bias.constant\_linear\_bias}). No redshift space distortions are taken into account.

\item [\texttt{WeakLensing}]
    \begin{multline}
 K_i(z) = \left( \frac{3 H_0^2\Omega_m}{2 c} \right) \left(\frac{(\ell+2)!}{(\ell-2)!} \right)^{1/2}\times 
 \\  (1+z)\ \chi(z) \int_z^\infty p_i(z^\prime)\ \frac{\chi(z^\prime)-\chi(z)}{\chi(z^\prime)}\ dz^\prime + K_{IA}(z)
    \end{multline}
where $K_{IA}(z)$ is an optional kernel function to deal with the Intrinsic Alignment. The implementation of this term currently follows \citet{2011A&A...527A..26J}, and is given by:
\begin{equation}
    K_{IA}(z) =   \left(\frac{(\ell+2)!}{(\ell-2)!}\right)^{1/2}\ p_i(z)\ b(z)\  H(z)\ \frac{C\  \Omega_m}{D(z)}
\end{equation}
with $C\approx 0.0134$ a dimensionless constant and $D(z)$ the growth factor of linear perturbations.
\end{description}

% Notice that all described kernel functions can be also viewed as function of the scale factor $a$ or as function of the radial comoving distance $\chi$.

% \FrL{We probably actually want to start by the Limber formula, which explicits what a kernel is, otherwise people don't necessarily know where these kernels come from.}

Because, like the other ingredients, the kernels are implemented as pyTree objects, \texttt{NumberCounts} and \texttt{WeakLensing}, all the integrals involved in these computations can be  differentiated with respect to the cosmological parameters  or the number densities using \textit{autodiff} and \jaxcosmo's implementation of integration quadrature. An example is given in the context of DES Y1 3x2pts analysis (Sec~\ref{sec-DESY1}).
%
\subsection{Validation against the Core Cosmology Library (CCL)}
%
To illustrate the different features available with the present version of the library (\jaxcosmo\ \texttt{0.1rc9}, which is available in the Python Package Index PyPI\footnote{\url{https://pypi.org/}}), we have written a  companion notebook \nblink{CCL_comparison} to compare it to the well-validated  Core Cosmology Library \citep{2019ApJS..242....2C}\footnote{\url{https://ccl.readthedocs.io}, version \texttt{2.5.1}.}. As examples, Figures \ref{fig:chi_comparison},
\ref{fig:halofit_comparison} and \ref{fig:Cell_comparison}
show the radial comoving distance (Eq.~\ref{eq:radial_comoving}), the non-linear matter power spectrum computation, and the angular power spectrum for galaxy-galaxy lensing (Eq.~\ref{eq:Cell_limber}) using the \texttt{NumberCounts} and \texttt{WeakLensing} kernel functions. \jaxcosmo\ features a suite of validation tests against CCL, automatically validating the precision of all computations to within the desired numberical accuracy; the relative differences between the two libraries are at the level of few $10^{-3}$ or better.

These numerical differences are mostly due to different choices of integration methods and accuracy parameters (eg. quadrature number of points). Increasing these parameters lead to performance degradation for \jaxcosmo\ as it increases the XLA compilation memory requirements significantly, especially for the angular power spectra computation. Since these differences are likely to be within the tolerance of the current generation of cosmological surveys, this trade-off is an acceptable one.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/comp_radial_com_dist.png}
    \caption{Comparison of the radial comoving distance between CCL and \jaxcosmo. More plots are available in the  companion notebook \nblink{CCL_comparison}.} 
    \label{fig:chi_comparison}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/halofit_pk.png}
    \caption{Comparison of the non-linear matter power spectrum (\textit{halofit} function) between CCL and \jaxcosmo. Is also shown the linear power spectrum.}     \label{fig:halofit_comparison}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/comp_Cell_shape_pos.png}
    \caption{Comparison of the shape-position angular power spectrum between CCL and \jaxcosmo.} 
    \label{fig:Cell_comparison}
\end{figure}

\section{Unleashing the Power of Fisher Forecasts with Automatic Differentiation}
\label{sec-fisher-forecast}

As a first illustration of the interest of \textit{autodiff} in cosmological computations, we present in this section a few direct applications involving the computation of the Fisher information matrix, widely used in cosmology. \\
Not only does the computation of the Fisher matrix become a trivial business, but the Fisher matrix itself becomes differentiable, allowing in particular for powerful survey optimization applications.

%JEC 5 August 22: not sure that should be given here as we have an Appendix for that and the main application described in the following sections concern DES Y1. 
%\subsection{DES Y1 setting}
% Here we describe the data and 3x2pt model for DES Y1 
% In this section, we adopt as a setting a DES Y1 3x2pt-like analysis.

\subsection{Instantaneous Fisher Forecasts}
Fisher matrices are a key tool in cosmology forecasting and experimental planning. By computing the Hessian matrix of a likelihood with respect to its parameters, we can find a Gaussian approximation to a posterior, which is usually sufficient for comparing the constraining power of different experimental configurations. As noted above, computing Fisher matrices is notoriously error-prone, since finite difference approximations to likelihoods must be carefully tuned for convergence, and observable calculations can be numerically unstable. Autodiff can help avoid this challenge.

We first illustrate, with an artificial case study, the computation of a Fisher matrix \citep{1997ApJ...480...22T,Stuart1991} using two methods with the \textit{autodiff} ability of JAX. For the detailed implementation, the reader is invited to look at the following companion notebook \nblink{Simple-Fisher}. In this example we use four tracer redshift distributions: two to define \texttt{WeakLensing} kernels and two for \texttt{NumberCounts} kernels. Then, the $10$ angular power spectra $C_\ell^{p,q}$ ($p,q:1,\dots,4$) are computed for $50$  logarithmically-spaced angular moments between $\ell=10$ and $\ell=1000$ using Equation \ref{eq:Cell_limber}. The Gaussian covariance matrix is computed simultaneously. A dataset is obtained from the computation of the $C_\ell^{p,q}$ with a fiducial cosmology. Then, the following snippet shows the log likelihood function $\mathcal{L}(\theta)$ implementation considering a constant covariance matrix ($\theta$ stands for the set of cosmological parameters). 
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
@jax.jit
def likelihood(p):
    # Create a new cosmology at these parameters
    cosmo = jc.Planck15(Omega_c=p[0], sigma8=p[1])
    # Compute mean and covariance of angular Cls
    mu, C = jc.angular_cl.gaussian_cl_covariance_and_mean(cosmo, ell, tracers, sparse=True)
    # Return likelihood value assuming constant covariance, so we stop the gradient
    # at the level of the precision matrix, and we will not include the logdet term
    # in the likelihood
    P = jc.sparse.inv(jax.lax.stop_gradient(C))
    r = data - mu
    return -0.5 * r.T @ jc.sparse.sparse_dot_vec(P, r)
\end{lstlisting}
%\end{minted}
The \texttt{jc.sparse} functions are implementation of block matrix computations: a sparse matrix is represented as a 3D array of shape $(n_y, n_x, n_{diag})$ composed of $n_y \times n_x$ square blocks of size $n_{diag} \times n_{diag}$.  The \texttt{jax.jit} decorator builds a compiled version of the function at the first use.

The first approach to obtaining approximate 1-sigma contours of the two parameters ($\Omega_c, \sigma_8$) with aa Fisher matrix uses the Hessian of the log-likelihood as follows:
\begin{equation}
F_{i,j} = - \frac{\partial^2\mathcal{L}(\theta)}{\partial \theta_i \partial \theta_j}
\qquad (\theta_1=\Omega_c, \theta_2=\sigma_8)
\label{eq:fisher_way1}
\end{equation}
which is accomplished in two lines of JAX code:
\begin{lstlisting}[language=iPython]
hessian_loglike = jax.jit(jax.hessian(likelihood))
F = - hessian_loglike(params)
\end{lstlisting}

The second approach to computing the Fisher matrix, which is restricted to Gaussian likelihoods but is more commonly used in the field because of its numerical stability, is to define a function that computes the observable mean  $\mu(\ell; \theta)$; it follows that the Fisher matrix elements are
\begin{equation}
    F_{i,j} = \sum_\ell \frac{\partial \mu(\ell)}{\partial \theta_i}^T C^{-1}(\ell)\frac{\partial \mu(\ell)}{\partial \theta_j}
    \label{eq:fisher_way2}
\end{equation}
where $C^{-1}(\ell)$ is the covariance matrix computed with the fiducial cosmology. This can also be simply computed in JAX: 
\begin{lstlisting}[language=iPython]
# We define a parameter dependent function that computes the mean
@jax.jit
def jc_mean(p):
    cosmo = jc.Planck15(Omega_c=p[0], sigma8=p[1])
    # Compute signal vector
    mu = jc.angular_cl.angular_cl(cosmo, ell, tracers)
    # We want mu in 1d to operate against the covariance matrix
    return mu.flatten() 
# We compute it's jacobian with JAX, and we JIT it for efficiency
jac_mean = jax.jit(jax.jacfwd(jc_mean))
# We can now evaluate the jacobian at the fiducial cosmology
dmu = jac_mean(params)
# Now we can compose the Fisher matrix
F = jc.sparse.dot(dmu.T, jc.sparse.inv(cov), dmu)
\end{lstlisting}
% done by JEC 13Sept22 \JZ{dmu in the above is not defined anywhere}

JAX implementations of the two methods agree to near perfect accuracy, as shown in Figure \ref{fig:simple_fisher_1}.
\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/simple_fisher_1.png}
    \caption{Comparison of the two methods to compute the Fisher matrix (Eqs.\ref{eq:fisher_way1},\ref{eq:fisher_way2}).} 
    \label{fig:simple_fisher_1}
\end{figure}
%
It is worth noting that in the two methods for Fisher matrix computations described above, the user does not need to vary individual parameter values to compute the 1st or 2nd order derivatives; this is in contrast to the usual finite difference methods. As an illustration, we have used the CCL library to compute the Fisher matrix via Equation \ref{eq:fisher_way2}. To do so, the Jacobian ($\partial \mu/\partial\theta_\alpha$) is computed with centered  order 4 finite differences available in the \texttt{Numdifftools} Python package\footnote{\url{https://numdifftools.readthedocs.io}}. Using the parameter values spaced by ($10^{-6}$, $10^{-2}$, $10^{-1}$) one obtains three different approximation of the 1-sigma contours as shown on Figure \ref{fig:simple_fisher_2}. The contour that agrees best with the \jaxcosmo\ method is obtained with the intermediate spacing parameter $10^{-2}$, implying that the user must tune this parameter carefully. Although very simple, this case study exhibits a significant challenge of finite difference methods for computing the Fisher matrix as has been shown for instance in a more advanced case study in \citet{2021arXiv210100298B}. 
% done by JEC 13Sept22 \JZ{Should probably latex the step sizes $10^{-1}, 10^{-2}, 10^{-6}$ in the figure legend}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/simple_fisher_2.png}
    \caption{Fisher matrices (Eq.\ref{eq:fisher_way1}) estimated using the CCL for the angular power spectra estimation and finite difference method to get the Jacobian  with different spacing of parameters ($10^{-6}$, $10^{-2}$, $10^{-1}$). For comparison, the contour obtained with \jaxcosmo\ is reproduced from Figure \ref{fig:simple_fisher_1} in black.} 
    \label{fig:simple_fisher_2}
\end{figure}

%
\subsection{Survey Optimization by FoM Maximization}
\label{sec-FOM-optimisation}
% Needs to mention the tomo challenge, mention that we can backpropagate through a NN and the cosmology model
% Maybe this can be only mentioned in the discussion.
\FrL{Insist here on the fact that the FoM itself is differentiable.}
Fisher forecasts are also commonly used in survey and analysis strategy strategy, where running a full posterior analysis for each possible choice would be unfeasible. The inverse area of a projection of a Fisher matrix in parameters of interest can be used as a metric for survey constraining power, such as in the Dark Energy Task Force report \citep{2006astro.ph..9591A}.

\jaxcosmo\ was used in a recent example of such a process, for the  the LSST-DESC 3x2pt tomography optimization challenge \citep{2021OJAp....4E..13Z}, where the best methodology for assignment of galaxies to tomographic bins was assessed using several such figures of merit and related calculations. The \jaxcosmo\ metric proved to be stable and fast.

\subsection{Massive Optimal Compression in 3 Lines}

\FrL{Here we can implement MOPED for the compression}

Once the Fisher matrix can be safely estimated, one can use the MOPED\footnote{Massively Optimised Parameter Estimation and Data compres-
sion} algorithm \citep{2000MNRAS.317..965H,2016PhRvD..93h3525Z, 2017MNRAS.472.4244H}. In the case of the constant covariance matrix the algorithm compresses data in a way that is lossless at the Fisher matrix level (i.e. Fisher matrices estimated using the compressed and full data are identical, by construction)  which reduces a possibly large data vector $\mu$ of size $N$ to $M$ numbers, where $M$ is the number of parameters $\theta_i$ considered. For instance, in the previous section, $N=500$ as $\mu=(C_\ell^{p,q})$ and $M=2$ for $(\Omega_c,\sigma_8)$. The algorithm computes by iteration $M$ vectors of size $N$ such that (taking the notation of the previous section)
\begin{equation}
    b_i = \frac{C^{-1}\mu_{,i}-\sum_{j=1}^{i-1}(\mu_{,j}^T b_j)b_j}{\sqrt{F_{i,i}-\sum_{j=1}^{i-1}(\mu_{,j}^T b_j)^2}}
\end{equation}
where $\mu_{,i}=\partial \mu/\partial \theta_i$ ($i=1,\dots,M$). The vectors $(b_i)_{i\leq M}$ satisfy the following orthogonality constraint
\begin{equation}
    b_i^T C b_j = \delta_{i,j}
\end{equation}
The algorithm is similar to the Gram-Schmidt process, using the constant covariance matrix $C$ to define the scalar product $\langle b_i, b_j\rangle$. Then, the original data set $x=(C_\ell^{p,q})_{p,q,\ell}$ is compressed in a data set composed of $M$ numbers $(y_i)_{i\leq M}$ according to
\begin{equation}
    y_i = b_i^T x
\end{equation}
These numbers are uncorrelated and of unit variance and this construction ensures that the log-likelihood of $(y_i)_{i\leq M}$  $(\theta_i)_{i\leq M}$, $y_i$ is identical to that of $x$ up to second order, meaning that the Fisher matrices derived from the two parameters should be identical, and in general the $y$ values should lose very little information compared to the full likelihood.

In problems where a (constant) covariance matrix is estimated from simulations, the number of such simulations required for a given accuracy typically scales with the number of data points used. 
MOPED therefore greatly reduces this number, often by a factor of hundreds. Since the uncertainty in the covariance due to a finite number of simulations must be accounted for \citep{2018MNRAS.473.2355S,2007A&A...464..399H}, this reduction can also offset any loss of information from the compression. \FrL{hummm but we still do need a full covariance in the compression algorithm, right?} \FrL{And we still do need to invert it... is the idea rather that if we dont have a perfect matrix at that stage the worse that can happen is sub optimal compression?}
%JEC{Well there are several ways that $C$ can be questioned: parameter dependence (ie. fiducial model/true model), non-gaussianities, errors... So not sure that all these effects induce only sub-optimal compression. But it may be out of the scope of this paragraph/paper.}

Another key advantage of the MOPED algorithm is to eliminate the need of large covariance matrix inversion of size $N\times N$ requiring $O(N^3)$ operations. This inversion takes place not only for the Fisher matrix computation (Eq.~\ref{eq:fisher_way2}, but more importantly in the log-likelihood computation (see the snippet of the previous section). The MOPED algorithm reduces the complexity to $O(M)$ operations.

To give an illustration, the following snippet uses the mock data set and results on the Fisher matrix computation from the previous section, to obtain the MOPED compressed data set composed of two parameters $(y_0,y_1)$ with maximal information on $(\Omega_c, \sigma_8)$:  
\begin{lstlisting}[language=iPython]
# Orthogonal vectors
b0 = jc.sparse.dot(C_inv,dmu[:,0])/jax.numpy.sqrt(F[0,0])
a = dmu[:,1].T @ b0
b1 = (jc.sparse.dot(C_inv,dmu[:,1]) - a * b0)/jax.numpy.sqrt(F[1,1]-a*a)
# MOPED vectors
y0 = b0.T @ data
y1 = b1.T @ data
\end{lstlisting}
Then, the log-likelihood can easily be implemented as:
\begin{lstlisting}[language=iPython]
@jax.jit
def compress_likelihood(p):
    # Create a new cosmology at these parameters
    cosmo = jc.Planck15(Omega_c=p[0], sigma8=p[1])
    # Compute mean Cl
    mu = jc.angular_cl.angular_cl(cosmo, ell, tracers).flatten()
    # likelihood using the MOPED vector
    return -0.5 * ((y0 - b0.T @ mu)**2 + (y1 - b1.T @ mu)**2)
\end{lstlisting}
The comparison between contour lines obtained with the original likelihood (uncompressed data set) and the MOPED version are shown in Figure \ref{fig:moped} in the case study of the previous section. Clearly, close to the negative likelihood minimum the lines agree very well.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/moded.png}
    \caption{Illustration of the log-likelihood contours obtained with an uncompressed data set (plain lines) and a MOPED version (dashed lines).} 
    \label{fig:moped}
\end{figure}
%
The case where the covariance matrix depends upon the parameters is discussed by  \citet{2017MNRAS.472.4244H} and leads to similar advantages of the MOPED compression algorithm.

%
\section{Posterior Inference made fast by Gradient-Based Inference Methods}
\label{sec:chmc}
%
In the following sections we review some statistical methods which directly benefit from the automatic differentiablity of \jaxcosmo likelihoods. We demonstrate a range of gradient-based methods, from Hamiltonian Monte Carlo (HMC), and its \textit{No-U-Turn} (NUTS) variant, to Stochastic Variational Inference (SVI). We further report on their respective computational costs in a DES-Y1 like analysis. All  the methods have been implemented using the \numpyro\ probabilistic programming language (PPL). 
%
% and NUTS after a \textit{Neural Transport} perform using a \textit{Stochastic Variational Inference} (aka SVI). All  the methods have been implemented using the \numpyro\ probabilistic programming language (PPL). 
%
%

\subsection{Description of the DES-Y1 exercise}
\label{sec-DESY1}
%JEC{Transfer of the model context from appendix here to ease the reading.}
From the DES server\footnote{\url{http://desdr-server.ncsa.illinois.edu/despublic/y1a1_files/chains/2pt_NG_mcal_1110.fit}} we have extracted the $N(z)$ distributions of the sources and normalize them to $[1.47, 1.46, 1.50, 0.73]$ effective number of sources per $\mathrm{arcmin}^2$ for each of the four redshift bins, respectively. The lens distributions $N(z)$ for five redshift bins are also extracted. The corresponding shapes of the different distributions are performed thanks to \texttt{jc.redshift.kde\_nz} function (``\texttt{jc}'' hereafter stands for \texttt{jax\_cosmo}) and are presented in Figure \ref{fig-DESY1-src-lens-redshift}.
\begin{figure}
\centering
\includegraphics[height=3cm]{figures/DESY1-source-redshift.png}
\includegraphics[height=3cm]{figures/DESY1-lens-redshift.png}
\caption{Distributions of the sources and lenses for the different redshift bins considered.}
\label{fig-DESY1-src-lens-redshift}
\end{figure}
\begin{table}[htb]
\caption{Priors of the 21 variables of the DES-Y1 of the 3x2pt likelihood (number counts and shear).}
\label{tab-DESY1}
 \centering
\begin{tabular}{ccccccccccc}
\hline
 parameter &  prior \\
 \hline
  \multicolumn{2}{c}{\textbf{Cosmology}} \\
$\Omega_c$ & $\mathcal{U}[0.10, 0.9]$ \\ %JEC 21/10/22 there was exchanges in Omega_c, sigma8, Omega_b intervalle (but ok in the notebooks)
$\sigma_8$ & $\mathcal{U}[0.40, 1.0]$ \\
$\Omega_b$ & $\mathcal{U}[0.03, 0.07]$ \\
$w_0$ & $\mathcal{U}[-2.00, -0.33]$ \\
$h$ & $\mathcal{U}[0.55, 0.91]$ \\
$n_s$ & $\mathcal{U}[0.87, 1.07]$ \\
\multicolumn{2}{c}{\textbf{Intrinsic Alignment}} \\
$A$ & $\mathcal{U}[-5,5]$ \\
$\eta$ &$\mathcal{U}[-5,5]$ \\
\multicolumn{2}{c}{\textbf{Lens Galaxy Bias}} \\
$(b_i)_{i:1,\dots,5}$ & $\mathcal{U}[0.8,3.0]$ \\
\multicolumn{2}{c}{\textbf{Shear Calibration Systematics}} \\
$(m_i)_{i:1,\dots,4}$ & $\mathcal{N}[0.012,0.023]$ \\
\multicolumn{2}{c}{\textbf{Source photo-$z$ shift}} \\
$dz_1$ & $\mathcal{N}[0.001,0.016]$ \\ % JEC note the sign is the one we have used but is different from https://arxiv.org/pdf/1708.01530.pdf
$dz_2$ & $\mathcal{N}[-0.019,0.013]$ \\
$dz_3$ & $\mathcal{N}[0.009,0.011]$ \\
$dz_4$ & $\mathcal{N}[-0.018,0.022]$ \\
\hline
\end{tabular}
\end{table}

Then, using the \numpyro\ PPL, we have set up a \textit{forward model} following the reference \citep{2018PhRvD..98d3526A}. The prior descriptions of the different model parameters are described in Table \ref{tab-DESY1}.  For instance the $\Omega_c$ parameter is treated as random variable as followed
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
Omega_c = numpyro.sample('Omega_c', Uniform(0.1, 0.9))
\end{lstlisting}
%\end{minted}

Then, we generate mock angular power spectra $C_\ell$ from the auto \& cross correlations of the Number Counts and Weak Lensing probes computed with the \texttt{gaussian\_cl\_covariance\_and\_mean} function in the \texttt{jc.angular\_cl} module. One reads
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
# Define the lensing and number counts probe
probes = [jc.probes.WeakLensing(nzs_s_sys, 
                       ia_bias=b_ia,
                       multiplicative_bias=m),
          jc.probes.NumberCounts(nzs_l, b)]
cl, C = gaussian_cl_covariance_and_mean(cosmo, 
                 ell, probes, 
                 f_sky=0.25, sparse=True)
\end{lstlisting}
%\end{minted} 
with \texttt{cosmo} an instance of the \texttt{jc.Cosmology} setting the cosmological parameters generated with the priors, and \texttt{ell} (ie. $\ell$) a series of 50 angular modes. The generation of the mock data (\texttt{cl\_obs}) yields from the possibility to execute a conditioned model with a fiducial cosmology, leaving \numpyro\ dealing with the random number generation:
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
fiducial_model = condition(model,
    {'Omega_c':0.2545, 'sigma8':0.801, 
    'h':0.682, 'Omega_b':0.0485, 'w0':-1.,'n_s':0.971,
     'A':0.5,'eta':0.,
     'm1':0.0,'m2':0.0,'m3':0.0,'m4':0.0,
     'dz1':0.0,'dz2':0.0,'dz3':0.0,'dz4':0.0,
     'b1':1.2,'b2':1.4,'b3':1.6,'b4':1.8,'b5':2.0
      })

with seed(rng_seed=42):
    cl_obs, P, C = fiducial_model()
\end{lstlisting}
%\end{minted}

The \textit{inference model} looks very similar to the \textit{forward model} excepts that first we fix the angular power spectra covariance matrix $C$ (and its inverse $P$) and secondly we let \numpyro\ generates the model $C_{\ell}$ from the priors
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
cl = jc.angular_cl.angular_cl(cosmo, ell, 
                              probes).flatten()
return numpyro.sample('cl', MultivariateNormal(cl, 
                            precision_matrix=P,
                            covariance_matrix=C))
\end{lstlisting}
%\end{minted}

These theoretical $C_{\ell}$ are in turn conditioned by the mock $C_\ell$ as followed
%\begin{minted}[fontsize=\footnotesize]{python}
\begin{lstlisting}[language=iPython]
observed_model = condition(model, {'cl': data})
\end{lstlisting}
%\end{minted}
%Then, as an example the optimisation of the cosmological parameters using for instance the NUTS flavour of MCMC can be conducted as followed
%\begin{minted}[fontsize=\footnotesize]{python}
%\begin{lstlisting}[language=iPython]
%nuts_kernel = numpyro.infer.NUTS(observed_model,
%              step_size=1e-1, 
%              init_strategy=numpyro.infer.init_to_median,
%              dense_mass=True,
%              max_tree_depth=7)
%
%mcmc = numpyro.infer.MCMC(nuts_kernel, 
%                          num_warmup=200, 
%                          num_samples=1000,
%                          num_chains=16,
%                          chain_method='vectorized',
%                          progress_bar=False)
%
%mcmc.run(jax.random.PRNGKey(42))
%\end{lstlisting}
%%\end{minted}
%More details can be found in the repository of this paper to perform a \numpyro\ modeling using \jaxcosmo: a structured NUTS sampling described in Section~\ref{sec-NUTS} \nblink{VectorizedNumPyro_block}, and a SVI optimisation followed by NUTS sampling of Section~\ref{sec-Neural-Reparametrisation} \nblink{DESY_Y1_SVI_and_NeutraReparam}. 
How to use this model to perform inference is described in the sections \ref{sec-NUTS} and \ref{sec-SVI}.
%
\subsection{Vanilla Hamiltonian Monte Carlo}
%
% Show Joe's vanilla HMC results against Cobaya
Hamiltonian Monte Carlo (HMC) is an MCMC-type method particularly suited to drawing
samples from high dimensional parameter spaces.  It was introduced in \citep{1987PhLB..195..216D}
and developed extensively since.  See \citet{betancourt} for a full review; we describe
very basic features here.

HMC samples spaces by generating particle trajectories through the space, using the log-posterior as the negative potential energy of a particle at each point $q$ in the space. Associated to $q$, we introduce an auxiliary $p$ variable in the spirit of Hamiltonian momentum such that
\begin{equation}
- \log{\cal P}(q) = V(q) + cte; \quad H(q,p) = V(q) + U(p)
\end{equation}
where $U(p)$ is a sort of \textit{kinetic} term defined by 
\begin{equation}
U(p) = p^T M^{-1} p
\end{equation}
At each sample, a trajectory is initialized with a random momentum $p$, and then integrated using Hamilton's equations with a numerical integration :
\begin{align}
\frac{\mathrm{d}p}{\mathrm{d}t} &= - \frac{\partial V}{\mathrm{d} q} = \frac{\partial \log{\cal P}}{\mathrm{d} q} \\
\frac{\mathrm{d}q}{\mathrm{d}t} &= + \frac{\partial U}{\mathrm{d} p} = M^{-1} p
\end{align}
where $M$ is a mass matrix which should be set to approximate the covariance of the posterior. This is also used to set the scale of the random initial velocities. These differential equations may be integrated iterating over $L$ steps the following \textit{leapfrog} algorithm
\begin{align}
    p_{n+\frac{1}{2}} &= p_n -\frac{\varepsilon}{2} \frac{\partial V}{\mathrm{d} q}(q_n) \\
    q_{n+1} & = q_n +\varepsilon M^{-1} p_{n+\frac{1}{2}} \\
    p_{n+1} &=  p_{n+\frac{1}{2}} -\frac{\varepsilon}{2} \frac{\partial V}{\mathrm{d} q}(q_{n+1})
\end{align}
where $\varepsilon$ is a step size parameter.

Formally, the set of $n_\mathrm{dim}$ moment are treated as new parameters, and after 
some number of integration steps a final point in the trajectory is compared to the initial one,
and a Metropolis-Hastings acceptance criterion on the total energy $H(q,p)$ is applied. If the trajectory is perfectly simulation then this acceptance is unity, since energy is conserved; applying it allows
a relaxation to the integration accuracy.

That gradients $\partial \log{\cal P} / \mathrm{d} q$ can be estimated using finite differencing,
but this typically requires at least $4 n_{\mathrm{dim}} + 1$ posterior evaluations per point, greatly slowing it
in high dimension, and as with the Fisher forecasting is highly prone to numerical error. Automatically
calculating a Hessian, as in \jaxcosmo, makes it feasible and efficient. \FrL{I don't think it uses a Hesssian? you only need the derivative of the log posterior wrt parameters. The hessian would be needed to get a good estimate of M.}

\FrL{Needs a little paragraph to explain with HMC is supposed to be better than plain Metropolis-Hastings.}
Metropolis-Hastings suffers from random walk to explore the geometry of the target distribution to get the region of maximum probability (ie. typical set) which is particularly  penalising in large dimensions. The HMC dynamics allow to make large jumps to explore the region around a sample to guess the next good one.

The tricky point is that the \textit{leapfrog} algorithm needs tuning to set the number of steps as well as the step size of integration. The next section addresses this problem thanks to the No-U-Turn HMC version.

%Below we show an example set of HMC constraints using \jaxcosmo\ to constrain a simulated
%Dark Energy Survey Year One (DES-Y1) lensing and clustering-like (3x2pt) likelihood, using 21 parameters\footnote{We omit the 
%five clustering sample photometric redshift parameters}.  We re-parametrize the space to unit mass using the Cholesky 
%decomposition
%of the covariance as estimated with an initial Fisher matrix; this lets us reflect the trajectories when they reach the
%edge of the parameter space, as recommended in \citep{NIPS2015_8303a79b}.  We use a fixed number (25) of integration steps and 
%a fixed integration period (0.02), this was chosen heuristically to give a reasonable acceptance rate, but is likely to be
%far from optimal.  We compare to the highly-optimized \texttt{Cobaya} 
%Metropolis-Hastings implementation  %\citep{cobaya2,cobaya1} JEC it may preferable to have a single up to date reference in JCAP
%\citep{2019ascl.soft10019T,2021JCAP...05..057T}, initializing it with the Fisher matrix as a proposal and
%tuning thereafter.  We ran two processes for each sampler, with four CPU threads per process.
%
%The Hamiltonian sampler required around half the number of samples as the Metropolis (70,000 vs 140,000) to reach approximately the same
%convergence, as tested by the Gelman-Rubin test \citep{gelman}, though each sample was itself a few times slower
%on a CPU.  This proof-of-principle example can certainly be hugely improved using various extensions to HMC such as No U-Turn Sampling \citep{nuts} or sampling from the entire trajectories \cite{betancourt}.  On next generation GPU high-performance systems, the advantage will be large.
%
%\begin{figure}
%\includegraphics[width=0.95\columnwidth]{figures/hmc.pdf}
%\caption{Constraints on three of the 21 parameters of a simulated DES-Y1 3x2pt likelihood, using a simple Hamiltonian
%Monte Carlo sampler with \jaxcosmo\ (blue), compared to the \texttt{Cobaya} Metropolis-Hastings sampler (orange).}
%\end{figure}
%
%
%We note that a full Boltzmann code accurate enough to calculate the CMB spectra to Planck \citep{planck18}
%or future \citep{s4,simons} accuracy is currently beyond the scope of this project.  
%If this is needed then HMC could use a fast approximation to a CMB theory spectrum, and importance sampling applied to the complete chain.

\subsection{NUTS}
\label{sec-NUTS}
%
% Show the advantage of using NUTS 
% look at difference in efficiency in terms of how many times we need to call the model.

The No-U-Turn (\textit{NUTS})  variant of the traditional HMC sampler has been introduced in \citet{10.5555/2627435.2638586}. It aims to help finding a new point  $x_{i+1}$ from  the current $x_i$. Namely, the root algorithm of HMC deals with the \textit{leapfrog} integration which requires to set two main parameters: $\varepsilon$ the step size and $L$ the number of steps. As stated by the authors, NUTS  uses a recursive algorithm to build a set of likely candidate samples that spans a wide swath of the target distribution, stopping automatically when it starts to double back and retrace its steps. So, in an automatic manner the parameter $L$ is adjusted to produce statistically  a new point $x_{i+1}$ independent from $x_i$, at the expense of the increase the number of model calls. The user has to set a new  parameter  (\texttt{max\_tree\_depth}) which gives in power of $2$ the maximal number of model calls at each generation.

Both sampler HMC and NUTS are available in the \numpyro\ library. After the forward model creation for the DES-Y1 3x2pts exercise described in section \ref{sec-DESY1}: \begin{itemize}
    \item we have first applyied  an affine transformation on cosmological, intrinsic alignment and biais parameters to use an unique uniform prior $\mathcal{U}[-5,5]$ (Table~\ref{tab-DESY1}); %\FrL{these are not latent variables, but model parameters. Also people don't know what this reparametrisation means.}
    \item  %\FrL{People don't know what the warmup phase is...}
    we use a structured mass matrix consisting on essentially a diagonal form  with a block structure for the following sets of parameters $(\Omega_b,\Omega_c,\sigma_8,w_0,h)$ and $(b_i)_{i:1\dots5}$. This matrix structuring is motivated by the expected degree of correlation of the concerned parameters as can be shown for instance in Figure \ref{fig_cobaya_NUTS_SVI_bis}.
\end{itemize}

As an example the optimisation of the cosmological parameters using for instance the NUTS flavour of MCMC can be conducted as followed with \numpyro:
\begin{lstlisting}[language=iPython]
%nuts_kernel = numpyro.infer.NUTS(observed_model,
              step_size=1e-1,
              init_strategy=numpyro.infer.init_to_median,
              dense_mass=True,
              max_tree_depth=7)

mcmc = numpyro.infer.MCMC(nuts_kernel, 
                          num_warmup=200, 
                          num_samples=1000,
                          num_chains=16,
                          chain_method='vectorized',
                          progress_bar=False)

mcmc.run(jax.random.PRNGKey(42))
\end{lstlisting}

The NUTS sampler have been running with 16 chains of 1,000 samples each after a warm-up phase consisting of 200 samples, and the \texttt{max\_tree\_depth} has been set to 7 by default (ie. 128 steps for each iteration). %\FrL{it should explain that vectorized means 16 chains in parallel on a single GPU.}
The "vectorized" option runs multi chains in a single GPU thanks to the JAX \textit{vmap} mechanism. If one dispose of several GPU devices, then the using the JAX paralellization mechanism (\textit{pmap}), it is possible to launch on each GPU device the "vectorized" sampling as done in the previous snippet, and get back all the MCMC chains. However, experiments have been undertaken on a single GPU either a NVidia Titan Xp (12GB RAM) on a desktop or an NVidia V100 (32GB RAM) at the IN2P3 Computing Centre\footnote{\url{https://cc.in2p3.fr/en/}}. The elapse time for these experiments was 20 hours.

The results in terms of relative effective sample sizes (aka ``ESS'') are detailed in Table~\ref{tab-ESS-NUTS_SVI-1} while the confidential level (CL) contours are presented in Figure \ref{fig_cobaya_NUTS_SVI}. One notices the dramatic improvement of the ESS by about a factor 10 using the NUTS sampler compared to the highly-optimized \texttt{Cobaya} Metropolis-Hastings implementation \citep{2019ascl.soft10019T,2021JCAP...05..057T} which has last about 40 hours on CPU to generate 70,000 samples. The CL contours are also in very good agreement. \FrL{we should rephrase this a bit to say that Cobaya is our reference.} It is worth to point out that the mass matrix structuring operation is a very effective as it increases the sampling efficiency by about a factor 2.   Anticipating the Section~\ref{sec-Neural-Reparametrisation} devoted to the Neural Transport which is an effective solution to validate the model with fewer samples, we have used a NUTS sampling with \texttt{max\_tree\_depth=5} those elapse time is 5 hours. This clearly shows a linear scaling with respect to the number of steps per iteration keeping the sampling efficiencies at a high level (see Table~\ref{tab-ESS-NUTS_SVI-1}), then the user is highly encouraged to tune this critical parameter. 
%
\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/fig_Cobaya-NUTS-SVI200_v1.pdf}
\caption{Constraints (30\%, 68\%, 90\%) on 4 of the 21 parameters of a simulated DES-Y1 3x2pt likelihood, using the \texttt{Cobaya} Metropolis-Hastings sampler (full curve in blue, 70,000 samples), the  NUTS sampling described in section \ref{sec-NUTS} (dashed curve in red; 16,000 samples), and a SVI MVN optimisation followed by a Neural Transport parametrisation to run a NUTS sampling (dashed-dotted curve in green, 200 samples).}
\label{fig_cobaya_NUTS_SVI}
\end{figure*}

\subsection{Stochastic Variational Inference}
\label{sec-SVI}
%
Let us remind briefly what is the \textit{Stochastic Variational Inference} method \citep{10.5555/2567709.2502622, 8588399} and show where auto-differentiation helps. If one  writes $p(z)$ the prior, $p(\mathcal{D}|z)$ the likelihood and $p(\mathcal{D})$ the marginal likelihood, then thanks to the  Bayes theorem one gets $p(z|\mathcal{D})=p(z)p(\mathcal{D}|z)/p(\mathcal{D})$ the posterior distribution of a model with latent variables $z$ and a set of observations $\mathcal{D}$. VI aims to get an approximation of this distribution, ie. $p(z|\mathcal{D}) \approx q(z;\lambda)$, by determining the variational parameters $\lambda$ of a predefined distribution. To do so, one uses the Kullback-Leibler divergence of the two distributions $KL(q(z;\lambda)||p(z|\mathcal{D}))$ leading to the following relation
\begin{align}
\log p(\mathcal{D}) &= \mathtt{ELBO} +  KL(q(z;\lambda)||p(z|\mathcal{D})) \label{eq-ELBO} \\
\mathrm{with} \ \mathtt{ELBO} &\equiv -\mathbb{E}_{q(z;\lambda)}\left[ \log q(z;\lambda)\right] + \mathbb{E}_{q(z;\lambda)}\left[ \log p(z,\mathcal{D}) \right] 
\end{align}
which defines the \textit{evidence lower bound} (aka ELBO) that one aims to maximize to get the $\lambda$ values. So, the optimal variational distribution satisfies
\begin{equation}
q(z;\lambda^\ast) = \underset{q(z;\lambda)}{\mathrm{argmax}}\  \mathtt{ELBO} = 
\underset{\lambda}{\mathrm{argmin}}\ \mathcal{L}(\lambda)
\end{equation}
The function $\mathcal{L}(\lambda)$ is the cost function used in practice. It is composed of two parts:
\begin{equation}
\mathcal{L}(\lambda) = \underbrace{\mathbb{E}_{q(z;\lambda)}\left[ \log q(z;\lambda)\right]}_{guide} - \underbrace{\mathbb{E}_{q(z;\lambda)}\left[ \log p(z,\mathcal{D}) \right]}_{model}
\label{eq-loss-svi-1}
\end{equation}
where the \textit{guide} in the \numpyro\ library may be a multi-variate Gaussian distribution (MVN) for instance.

Using the auto-differentation tool, one can use ``black-box'' guides (aka \textit{automatic differentiation variational inference}). As stated by the authors of \citep{10.5555/3122009.3122023} ADVI helps specifying a variational family appropriate to the model, computing the corresponding objective
function, taking derivatives, and running a gradient-based or coordinate-ascent optimization. First is defined a invertible differentiable transformation $T$ of the original latent variables $z$ into new variables $\xi$, such $\xi=T(z)$ and $z=T^{-1}(\xi)$. The main advantage of this reparametrisation is that the minimization can be performed with no bound constraints. One can develop ``AutoGuides'' (\numpyro\ terminology) that can be adapted to the user models. Then, the cost function reads
\begin{equation}
\mathcal{L}(\lambda) = \underbrace{\mathbb{E}_{q(\xi;\lambda)}\left[ \log q(\xi;\lambda)\right]}_{guide} - \underbrace{\mathbb{E}_{q(\xi;\lambda)}\left[ \log p(\xi,\mathcal{D}) \right]}_{model}
\label{eq-loss-svi-2}
\end{equation}
with
\begin{equation}
p(\xi,\mathcal{D}) \bydef p(T^{-1}(\xi),\mathcal{D}) |J_{T^{-1}}(\xi)|
\end{equation}
which exhibit the Jacobian of the $T^{-1}$ transformation.
The evaluation of the expectations during gradient descent of the loss Eq.~\ref{eq-loss-svi-2} can be done using what is called \textit{elliptical standardisation} or \textit{re-parametrization trick} or \textit{coordinate transformation} (see references in \cite{10.5555/2969239.2969303}).  Let us illustrated the method using an invertible transformation (aka \textit{normalizing flow})  $S_\lambda$ such that $S_\lambda(\xi)=\zeta$, such that $\zeta\sim \mathcal{N}(0,I)$\footnote{Notice that $z=T^{-1}(\xi)=(T^{-1} \circ S_\lambda^{-1})(\zeta)=(S_\lambda \circ T)^{-1}(\zeta)=F_\lambda(\zeta)$.}. The Jacobian of this distribution is 1 by definition (volume conservation), so the lost function reads
\begin{multline}
-\mathcal{L}(\lambda) = \underbrace{\mathbb{E}_{\zeta\sim \mathcal{N}(0,I)}\left[ \log p(T^{-1}(S_\lambda^{-1}(\zeta)),\mathcal{D}) + \log |J_{T^{-1}}(S_\lambda^{-1}(\zeta))| \right]}_{model} \\ + \underbrace{\mathbb{H}[q(\xi;\lambda)]}_{guide}
\label{eq-loss-svi-3}
\end{multline}
where $\mathbb{H}(q)$ is the Shannon entropy of the $q$ distribution, so its gradient can be computed once for all for a given $q$ distribution family and reused in any user model.  
Then,
to get $\nabla_\lambda \mathcal{L}$, the $\nabla$ operator can be put inside the expectation which leads to\footnote{To simplify the notation, $T^{-1}(S_\lambda^{-1}(\zeta))$ has been replaced by $z$.}
\begin{multline}
-\nabla_\lambda\mathcal{L}(\lambda) = \mathbb{E}_{\zeta\sim \mathcal{N}(0,I)}\left\{
\left[ \nabla_z \log p(z,\mathcal{D}) \times \nabla_\xi[T^{-1}(\xi)] \right. \right. \\
+ \left. \left. \nabla_\xi \log|J_{T^{-1}}(\xi)| \right] \times \nabla_\lambda S_\lambda^{-1}(\zeta)
\right\}
+ \nabla_\lambda \mathbb{H}[q(\xi;\lambda)]
\label{eq-loss-svi-4}
\end{multline}
An concrete example of the gradient formula is given in Appendix \ref{app-SVI_MVN} with the multivariate Gaussian distribution family used as the guide.
%
%\FrL{here it would be better to have a small statement on the limitations of VI, and why it's a good idea to then go to Neutra.} 
%\FrL{we don't have plots of just the VI?}\JEC{Good point: I have SVI alone with both MVN or Block Neural Autoregressive Flow. See new figure.}

Here is the simple way \numpyro\ deals with SVI initialisation and optimisation\footnote{See also listing in Appendix \ref{sec-DESY1}) where the \texttt{model} is described and the \texttt{cl\_obs} are the observed $C_\ell$.}
\begin{lstlisting}[language=iPython]
import numpyro.infer.autoguide as autoguide
from numpyro.infer import SVI, Trace_ELBO
from numpyro.optim import Adam
# choose a SVI guide 
guide = autoguide.AutoMultivariateNormal(model,
        init_loc_fn=numpyro.infer.init_to_median())
# SVI optimisation 
optimizer = numpyro.optim.Adam(1e-3)
svi = SVI(model_spl, guide,optimizer,loss=Trace_ELBO(num_particles=10))
svi_result = svi.run(jax.random.PRNGKey(0),20_000, cl_obs)
#To get samples from SVI approximated posterior 
samples = guide.sample_posterior(jax.random.PRNGKey(1), svi_result.params, sample_shape=(100_000,))
\end{lstlisting}
%
Once the optimisation is done on gets $z$ i.i.d. samples from the $q(z,\lambda^\ast)$ distribution  applying the inverse of the inverse of the $S_{\lambda^\ast}$ and $T$ transformations. Using the same DES-Y1 simulation as in previous section, we use both a MultiVariate Normal distribution (MVN) and a Block Nueral Autoregressive Flow (B-NAF) \citep{pmlr-v115-de-cao20a} as \textit{guides} to approximate the true posterior (\textit{model}). The reader can have a glance to the following notebook \nblink{DES_Y1_SVI_and_NeutraReparam}. Concerning the B-NAF architecture it is composed of  a single flow using a block autoregressive structure with 2 hidden layers of 8 units each. The SVI optimization has been performed with the Adam optimizer \citep{KingmaB14} and a learning rate set to $10^{-3}$. We have stopped the optimization after 20,000 (30,000) steps to ensure a stable ELBO loss convergence when using the B-NAF (MVN) guides, and no tuning of the optimizer has been undertaken as for instance a learning rate scheduling. It takes about 2 or 3 hours on the NVidia V100 GPU scaling according to the number of steps.  \FrL{Arf... 2 hours !}

In the figure \ref{fig_cobaya_SVIs}, we compare the contours obtained with \texttt{Cobaya} as in figure \ref{fig_cobaya_NUTS_SVI}) and those obtained with the MVN and B-NAF guided SVI.  As mentioned in \cite{NEURIPS2020_7cac11e2} one challenge with variational inference is assessing how close the variational approximation $q(z,\lambda^\ast)$ is to the true posterior distribution. It is not in the scope of this article to elaborate a statistical diagnostics, rather we show that both guided SVI exhibit rather similar contours and both estimate are close to the Cobaya posterior sampling. The difference is that these SVI approximate posteriors have been obtained in a short time and these posterior estimates can serve as front-end to a NUTS sampler as described in the next section. 
%
\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/fig_Cobaya_SVIs_v1.pdf}
\caption{Same configuration as in Figure \ref{fig_cobaya_NUTS_SVI} but using sampling of approximated posterior using SVI with MVN and B-NAF guides (see text) compared to \texttt{Cobaya} sampling.}
\label{fig_cobaya_SVIs}
\end{figure*}
%
\subsubsection{Neural Transport}
\label{sec-Neural-Reparametrisation}
%
If the SVI method can be used as is to get $z$ i.i.d. samples from the $q(z,\lambda^\ast)$ distribution as shown on the previous section, the \textit{Neural Transport MCMC} method \citep{Parno2018,2019arXiv190303704H} is an efficient way to boost HMC efficiency, especially in target distribution unfavourable geometry where for instance the leapfrog integration algorithm has to face squeezed join distributions for a subset of variables. From SVI, one obtains a first approximation of the target distribution and we show that the change-of-variable $z=F_\lambda(\zeta)$ (eg. $F_\lambda=T^{-1}\circ S^{-1}_\lambda$) is such that
\begin{equation}
q(z;\lambda) \rightarrow q(\zeta;\lambda) \bydef q(F_\lambda(\zeta)) |J_{F_\lambda}(\zeta)|
\end{equation}
which is evaluated with optimal $\lambda^\ast$ according to maximization of the ELBO. So, one can use a HMC sampler (eg. NUTS) based on $p(\zeta;\mathcal{D})$ distribution, initialized with $\zeta$ samples from $q(\zeta;\lambda^\ast)$, to get a Markov Chain of $N$ samples $(\zeta_i)_{i<N}$. Then, from the transformation  $z_i=F_{\lambda^\ast}(\zeta_i)$ one finally obtain a Marcov Chain with $(z_i)_{i<N}$ samples. 
\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/fig_Cobaya-NUTS-SVI200_bis_v1.pdf}
\caption{Same configuration as in Figure \ref{fig_cobaya_NUTS_SVI} but for the highly correlated five lens galaxy bias.}
\label{fig_cobaya_NUTS_SVI_bis}
\end{figure*}
The \textit{Neural Transport} is accessible throw the  \verb|NeuTraReparam| in \numpyro\ library as followed (See the notebook \nblink{DES_Y1_SVI_and_NeutraReparam}):
\begin{lstlisting}[language=iPython]
from numpyro.infer.reparam import NeuTraReparam
# Reparametrisation of model to feed NUTS sampler
neutra = NeuTraReparam(guide, svi_result.params)
neutra_model = neutra.reparam(model_spl) 
\end{lstlisting}
%
We have used the MVN guided SVI described in the previous section. The NUTS sampler has run 1 chain of 200 samples (``low'' schema), 1 chain with 1000 samples and 10 chains  with 1000 samples each (each setup starts all chains with 200 samples for initialisation). All NUTS sampling as been performed with dense mass matrix optimisation without special block structuring and with \texttt{max\_tree\_depth=5} which is different than the default NUTS setting described in section \ref{sec-NUTS}. The elapse time for each of the 3 setups is 50 minutes, 150 minutes and 5 hours, respectively. Naturally, more samples lead to better contour precision. But what is illustrative is the ``low'' schema results as shown in the Figures \ref{fig_cobaya_NUTS_SVI} and \ref{fig_cobaya_NUTS_SVI_bis}  to be compared to \texttt{Cobaya} and NUTS results presented in section \ref{sec-NUTS}. Notice the good results even for the highly correlated lens bias parameters. It is noticeable that running NUTS with the same settings  but without the SVI Neural Transport phase has demonstrated a rather poor behaviour with only 200 samples.

Results in terms of sampling efficiency are shown on Table~\ref{tab-ESS-NUTS_SVI-1} showing that the SVI followed by Neutral Transport gives high efficiency at low number of samples which is quite interesting to ease the model development in its early phase.
%
\subsection{Sampling efficiency}
\label{sec-results}
%
To a large extend looking at results of Table~\ref{tab-ESS-NUTS_SVI-1}, the key question is: what is the HMC/NUTS gain compared to the well known highly optimized \texttt{Cobaya} sampler? 
In one hand, one may argue considering the elapse time of execution that the number of efficient samples per model call is a right metric to consider: 
\begin{equation}
    \eta = \frac{n_{eff}}{\#\ of\ calls} = \frac{N_s \times \varepsilon}{N_s \times n_{steps}} = \frac{\varepsilon}{n_{steps}}
\end{equation}
with $N_s$ the total number of samples, $\varepsilon$ the effective sampler efficiency  and $n_{steps}$ the number of steps (calls) per sample generation. Then, for \texttt{Cobaya} one roughly gets $\varepsilon\approx 3\%$ with $n_{steps}=1$ while for the NUTS sampler $\varepsilon\approx 50\%$ but at the expense of $n_{steps}=2^5$ or more. At the end of the day after tuning generation parameters, the $\eta$ values for both methods would certainly be comparable (ie. about $1\%$). So, at first sight one may conclude that there is no gain to use HMC/NUTS compared to standard MCMC sampler.

On the other hand, the above metric is may be too crude to cover all aspects of the sample generation. One should probably have in mind that the low sampling efficiency of standard MCMC sampler requires mobilizing a large amount of resources to produce a sufficient large sample batch in a reasonable time scale, ie. several days on dedicated infrastructure. Moreover, the user is faced to the dilemma to perform or not a statistical sample selection post-processing as \textit{thinning} beyond the \textit{burn-in removal} which is a rather delicate procedure and time consuming too, needing how-know to be conducted correctly \citep{doi:10.1146/annurev-statistics-040220-091727, Owen2017}.

So, the authors think that it is worth considering methods like HMC/NUTS that by construction are more efficient. Moreover, using SVI with neural reparametrisation offers an effective way to undertake a progressive validation of the model with rather modest sample set (eg. starting with 200 samples) producing good enough marginal contours in few hours. In practice, this validation phase can save time before producing sizeable batch for final analysis. The authors have not investigated very high dimensional ($O(10^2)$ parameters) as well as multi-modal analysis, but the key argument in favour of HMC/NUTS sampling is that it exploit the geometry of the typical set of the posterior distribution contrary to the standard random walk of Metropolis-Hasting sampling. Furthermore, using reparametrisation  one can adapt to poor geometry cases (eg. \cite{2019arXiv190303704H}).
%
\begin{table}[htb]
\caption{Results of the \textit{relative effective sample size} (aka ESS, ``bulk'') in percent computed by the \texttt{Arviz} library \citep{arviz_2019}. The ESS can be larger than 100\% in some cases. 
%{\color{red} 
%(0) "F.L run" are computed from NUTS sampler with 16 %parallel chains of 1,000 samples each and 200 warm-up %and \texttt{max\_tree\_depth=7} (see text) The numbers %are from \texttt{n\_eff/16,000};
%}
(a): results from Cobaya 70,000 samples;
(b): results from NUTS sampler with 16 chains of 1,000 samples each and 200 warm-up and \texttt{max\_tree\_depth=7} (Section~\ref{sec-NUTS}); 
(c) are the results of SVI Multivariate Normal followed by NUTS and Neural Transform  with 1 chain of 200 samples and 200 warm-up (Section~\ref{sec-Neural-Reparametrisation}).}
\label{tab-ESS-NUTS_SVI-1}
 \centering
\begin{tabular}{ccccccccccc}
\hline
    & $\Omega_b$ & $\Omega_c$ & $\sigma_8$ & $w_0$ & $h$ & $n_s$ & $A$ & $\eta$\\
\hline
(a) &  $3.1$ & $2.5$       & $2.9$      & $2.9$  & $2.6$  & $3.1$  & $3.1$ & $2.8$ \\  
(b) & $48.1$ &  $45.6$     & $36.2$     & $33.4$ & $52.8$ & $50.1$ & $68.8$ & $48.8$\\
(c) & $84.0$ &  $28.0$     & $26.0$     & $20.5$ & $80.0$ & $110.5$ & $58.5$ & $29.5$\\
\hline
\end{tabular}
\end{table}
%
\section{General discussion}
\label{sec-discussion}
%
Having demonstrated the interest of \jaxcosmo\ as a differentiable cosmology library library, we now discuss several limitations, and open questions.

%Lack of autodiff Boltzmann code (emulator)
The first essential barrier to a full-fledged automatically differentiable cosmology library is the need for a differentiable Boltzmann solver to compute the matter power spectrum. At this stage, \jaxcosmo\ relies on the analytic Einsenstein \& Hu fitting formula, which is not accurate enough for Stage IV requirements, nor does it allow to explore beyond $\Lambda$CDM models. Existing solvers such as CLASS or CAMB \FrL{citations} are large and complex codes which are not easily reimplemented in an \autodiff\ framework and therefore cannot be directly integrated in \jaxcosmo\ .

A first option to resolve this issue is to implement from scratch a new Boltzmann code in a framework that will support automatic differentiation. This is the approach behind works such as \FrL{the Julia code} which is provides a simplified Boltzmann solver in Julia, or \FrL{the code from our friends in Zurich} which is based on a SymPy and could be relatively easily compatible with JAX. However, even if very promising, both of these options thus far remain limited. While we do believe a automatically differentiable Boltzmann code is the best option, it seems that the cost of developing such a code remains very high at this time.

A second approach is to develop emulators of a full-fledged Boltzmann code. Emulators based on neural networks or Gaussian processes are themselves automatically differentiable with respect to cosmological parameters. In fact, the litterature is now rich in examples of such emulators \FrL{cite all emulators}. After validating their accuracy against the reference CAMB or CLASS implementation, they can be directly integrated as a plug-and-play replacement for the computation of the matter power spectrum. \FrL{need to talk about cosmopower somewhere.} At this time, it seems that using emulators will be the most straightforward approach to bring more accurate models to \jaxcosmo\ . \FrL{However, we want to emphasize that one of the reason for the wide diversity of emulators is a lack of standardization}.

\bigskip

% Reasoning behind not making a full emulator of the correlation functions
Connected to this discussion about emulators, one point that could be discussed is whether one even needs a library like \jaxcosmo\ if anyway one can build emulators of a CCL likelihood for use inside gradient-based inference algorithms. While this could indeed be feasible, and of similar cost as just making an emulator of the matter power spectrum, the drawback of this approach is that many analysis parameters and choices become hard coded in the emulator. While the model for the linear matter power spectrum is typically kept fixed in practical analyses, all the of the choices related in particular to systematics modeling (e.g. photometric redshift errors, galaxy bias, or intrinsic galaxy alignments) are very likely to vary significantly in the process of developing the analysis. Building an emulator for the likelihood would mean the emulator needs to be retrained everytime the likelihood is changed.

\bigskip

%  - Massively Parallel MCMC on GPU (linked to vmap and/or pmap)
%  - Efficiency of gradient-based inference methods
Another aspect worth discussing are the prospects for scaling up and speeding up cosmological inference in practice given tools such as \jaxcosmo\ . As we illustrated in the previous section, gradient-based inference techniques yield significantly less correlated MCMC chains.\FrL{Hummm here I guess we want to say also it scales better with dimensions, and that you can get very fast first results.} \FrL{we also want to say that the GPU parallelisation is going to change the game, all the more when you account for new massive GPU machines.}

\bigskip

%  - Respective role of CCL and jax-cosmo
Finally, one question worth asking is how does \jaxcosmo\ position itself against classical codes such as CCL? Or in other terms, will it replace CCL? While we are convinced of the benefits of a JAX implementation, we expect CCL to remain extremely valuable as the reference cosmology implementation. \FrL{it allows to leverage the legacy codes, has a lot more features, etc.} Ultimately, a natural transition may occur towards more modern frameworks like \jaxcosmo\ when they reach the ability to run full-fledged Stage IV likelihoods.

\section{Conclusions \& Prospects}
\label{sec-conclusion}
%- We have presented a library using automatic differentiation and and automatic GPU offload for a set of key cosmology theory calculations
%- We currently cover only a small number of the many calculations needed in cosmo theory, and welcome contributions for other areas such as CMB.
%- The calculation of Fisher matrices using this method, which is infamous for requiring extensive careful tuning, becomes very simple and robust using JC.
%- We have demonstrated the efficiency in JC of the derivative-aware MCMC methods HMC and NUTS, which are regarded as the only way for samplers to evade the curse of dimensionality up to the hundreds of dimensions we are likely to need for the next generation of surveys.
%- We have shown a proof-of-concept for the use of JC with machine learning methods, opening a whole new space of methods in cosmology.

In this paper, we have presented \jaxcosmo\ a cosmology library implemented in the JAX framework, enabling the automatic computation of the derivatives of cosmological likelihoods with respect to their parameters, as well as greatly speeding up likelihood evaluations thanks to automatic parallelisation and just in time compilation on GPUs. Currently, \jaxcosmo\ only contains a small set of features limited corresponding to a DES-Y1 3x2pt analysis. Being an open source project, contribution of additional features is additional areas such as CMB or spectroscopic galaxy clustering are warmly welcome.

To demonstrate the interest of an automatically differentiable library, we have illustrated on concrete examples how Fisher matrices, which are notoriously difficult to compute and require extensive and careful fine tuning, can now be computed robustly and at much lower cost. In addition, Fisher matrices becomes themselves differentiable, which allows for  Figure of Merit optimization by gradient descent, making survey optimization extremely simple.

Going beyond Fisher forecasts, we have also compared to simple Metropolis-Hastings several gradient-based inference techniques (namely, Hamiltonian Monte-Carlo, No-U-Turn-Sampler, Stochastic Variational Inference). We show that the posterior samples with gradient-based methods have a much larger Effective Sample Size, meaning that \FrL{let's think a bit more about interpretation here.... the thing I care about is not as much the ESS, but the number of likelihood evaluation per independent sample. I don't think we have quantified that too much}. These inference techniques can scale to hundreds of dimensions, and may become necessary in Stage IV analysis as the number of nuisance parameters is likely to become very large.

\bigskip

The next extensions of this framework will be the inclusion of additional cosmological probes, as well as integrating emulators for the matter power spectrum trained on CAMB or CLASS, as a means to go beyond the current analytic Eisenstein \& Hu model.

\bigskip

In the spirit of reproducible research, all results presented in this paper can be reproduced from the following GitHub repository:

%\FrL{add link}\JEC{Verif if all the nb listed are there, and if some have to be withdrawn. }

\url{https://github.com/DifferentiableUniverseInitiative/jax-cosmo-paper/}

%
%
%
%\subsection{Efficient sampling in high dimensions with Hamiltonian Monte Carlo}
%
%%
%%
%\subsection{DES-Y1 Example}
%

%%%
\section*{CRediT authorship contribution statement}
\textbf{J.E Campagne} Conceptualization, Methodology, Software, Validation, Writing.
\textbf{F. Lanusse}, \textbf{J. Zuntz}...

%\section*{Declaration of Competing Interest}
%The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

%\Acknowledgements
\section*{Acknowledgements}
Some of the numerical experiments have been conducted at the IN2P3 Computing Center (CC-IN2P3 - Lyon/Villeurbanne - France) funded by the Centre National de la Recherche Scientifique.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix

\section{SVI example with MVN guide}
\label{app-SVI_MVN}
Let us illustrate the \textit{elliptical standardisation} described in section \ref{sec-SVI} in the case where $q(\xi; \lambda) = \mathcal{N}(\xi; \mu, \Sigma)$ (ie. a MVN distribution with the parameters $\lambda=(\mu, \Sigma)$). One defines $S_\lambda$ such that $S_\lambda(\xi)=\zeta$, eg. $\zeta = L^{-1}(\xi-\mu)$ where $L$ is the Cholesky decomposition of $\Sigma=LL^T$. The Jacobian of this distribution is 1 (volume conservation), so the cost function reads
\begin{multline}
-\mathcal{L}(\lambda) = \underbrace{\mathbb{E}_{\zeta\sim \mathcal{N}(0,I)}\left[ \log p(T^{-1}(S_\lambda^{-1}(\zeta)),\mathcal{D}) + \log |J_{T^{-1}}(S_\lambda^{-1}(\zeta))| \right]}_{model} \\ + \underbrace{\mathbb{H}[q(\xi;\lambda)]}_{guide}
\label{eq-loss-svi-3b}
\end{multline}
where $\mathbb{H}(q)$ is the Shannon entropy of the $q$ distribution which is independent of the model, so its gradient can be computed once for all and reused in any user model. In case of the MVN for completeness  $\nabla_\mu \mathbb{H}(q)=0$, $\nabla_L \mathbb{H}(q)=(L^{-1})^T$ while the other gradients are computed by the JAX \texttt{autodiff} mechanism.
%\FrL{Yeah. this is way TMI, and also goes against the idea that with autodiff we don't need to manually work out all these details.}

The expectations used in the above expressions are computed with $\zeta$ i.i.d samples from $\mathcal{N}(0,1)$ distribution which reveals the ``S'' of SVI. As one can imagine all the jacobian computations take benefit of the automatic differentiation offers by JAX.
%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%\section*{Bibliography}
%\bibliographystyle{elsarticle-harv} 
\bibliographystyle{aa} 
\typeout{}
\bibliography{refs}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.